\chapter*{Acknowledgments $\&$ Cariños}
\epigraphhead[10  ]{\epigraph{\textit{Gracias a la vida que me ha dado tanto }}{}}
Over all of the people I want to thank, John Calsamiglia (my PhD advisor) has undoubtedly been \textit{the} superhero. Among all the stuff he taught me ---which is certainly a lot by judging from the content of this thesis and our scientific collaborations---, there is one I want to highlight. Many issues, both scientific and life-related, have arisen during these years\footnote{As they say, \textit{La vida sin problemas es matar el tiempo a lo bobo.}}, and it was John who has taught me how to deal with such situations, and to find solutions without becoming desperate. These really subtle life lessons will mark me forever. I have no other words to thank this rather than saying that not only my scientific career but also personal life will be marked by this legacy. Per tot això, i tots aquests anys compartits, moltíssimes gràcies John!

Furthermore, this thesis would be really different if it wasn't for the two extraordinary post-docs I have worked with: Matteo Rosatti and Giulio Gasbarri. I also have no other words than big thanks to both of you, not only for the physics you taught me but also for the moments we enjoyed toghether.

During these years, the research group has been some sort of family to me, and it has been a pleasure to share the office with Matt, María, Minglai, Mani, Andreas, Anna, Ramón, Emili, Philipp, Gael, Suzana, Joe, Carlo, Naga, Illaria, Marco, Jennifer, Nikklas, Walther, Micha, Yago, Alonso, Jessica, Andreu, Antoni, Elisabeth, Esteban, Farzin, Zahra, and so many more!.

Tracing back a bit, I thank to all the people who encouraged me to move so far away from home to do this PhD. Above all, my family, who was an essential support during these times, even from so far away: mis viejos, mi hermana, Elena, Mariano, Flavia, Mandi, Paloma, mis abuelos Yaco, Scheine, Fanny y Abraham, Ruth, Tomi, Gonza, y todo el resto.

It was José Edelstein who first suggested me to move here (Barcelona); the Toni Acín, who has always been super receptive, was my first contact with this amazing city and got me in touch with GIQ. So big thanks to them!

Also, big thanks to all of my former teachers and collaborators from whom I have learnt and still learn so much; among them Mauricio Matera, Raúl Rossignoli, Norma Canosa, Lorena Rebón, Tomás Crosta, José Esté, Raul Morral, Marco Cerezo, Lukasz Cincio, Patrick Coles and Guillaume Verdon. Big thanks to the research institutions Instituto de Física La Plata and Los Alamos National Laboratory for hosting me during this years.

I also want to thank the unique support from people at Port d'Informació Científica (PIC), who lent me their computational resources which were essential to get most of the numerical results later analyzed in the thesis.

Also, I had an enormous help from my psychologist, who was essential to help me to take so many difficult decisions. Por el infinito apoyo durante estos años, muchas gracias Maren!

Son muchos años y por lo tanto mucha gente a la que agradecer, sin la cual toda esta aventura tampoco hubiera sido posible. Son al menos dos los lados del océano que toca agradecer, y las fronteras se suelen confundir épocas de tanta conexión. Sin más orden que el desorden, millones de gracias por todo el aguante a Valen Caso Rosendi, Caro Durián, Joaco Gil, Pedro Pablo, Javi Argüello, Mecchi Do Campo, Eric Kobrinsky, Amy Vianne, Sara Silva, Jero Penini, Seba Cabuli, Jose Giugno, Andi Sanelli, Ines Valla, Lucas Hadad, Nico Mártite, Nachito Piaggio (y toda la banda de natación platense!). Al maravishoso equipo Mashonesero: Cris Ceijas y Bea Astorga, a los vecinos de La Floresta y Vallvidrera: Carol Stankevicius, Ale Graziati, Taba Rios, Nati de la Fuente, Iña Allende, Lucho Griso y compañía. También Teresa Reinhard, Marta Solé Al Max Nordau en La Plata que siempre estará en mis raíces, a Lauti Krol, Pau Fioramonti, Eli Butinof, Juli Coloma, Joha Pronsky, Gasti Sal Anglada, Caro Bikis que es mi hermana pero ya que estamos la nombro de nuevo. ¡Cuánto aguante! También a Raulo Orcanza Rojas, Manu Hadad, Juli, Caro y Alicia Golijov (y Custom!). También Benja Ciancio, los Martines (Lagares y Tomé), Nahuel Díaz, Poe Enciso Porro, Pablo Gomez, Juani Sisti, Brian Evans, Floppy Espain, Sofi Lawrie, Manchi Machao, Nela Clinaz, Nacho Bruvera, Fiorela Ghilini, Cecchi, Lola Dafau-Saphores, Zenaida Andreica, Marta Fullola, A todo el equipo de triatló de la la Unió Ciclista de Sant Cugat! Cada persona que nombro me ha dado---y también las muchas que olvido nombrar---, durante estos años, un apoyo inigualable. Ah! and to my dog Dam that was always there (even virtually) for the routinary hugs :)
